---
title: "Machine Learning Course Project"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Project summary
Six participants were asked to perform barbell lifts correctly and incorrectly in 
five different ways. Use data from accelerometers on the belt, forearm, arm, and dumbell to predict the lift method.

http://web.archive.org/web/20161224072740/http:/groupware.les.inf.puc-rio.br/har 

1. Load trainig data
```{r}
library(data.table)
training <- fread("https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv")
```

2. Remove columns that do not add to the model to reduce size of dataset. Several
variables are derivatives of the other variables (max, min, avg, stddev, etc),
and mostly ( > 97%) NA. Remove columns with user name, time stamp, window

```{r message = FALSE, warning = FALSE}
library(dplyr)
library(caret)
remove <- c("name", "timestamp", "window", "max", "min", "kurtosis", "skewness", 
                  "amplitude", "var", "avg", "std", "V1")
train1 <- select(training, !contains(remove))
```

3. Check remaining predictors to see if any others can be removed
```{r}
nearZeroVar(train1, saveMetrics = FALSE)
```

outcome shows no additional removal or variables appropriate

4. Create a train and test subset of training dataset. This allows testing of 
accuracy of model without using the final test data
```{r}
inTrain <- createDataPartition(y = train1$classe, p = 0.70, list = FALSE)
subset_train <- train1[inTrain,]
subset_test <- train1[-inTrain,]

dim(subset_train)
dim(subset_test)
```

5. Fit a model using boost method (gbm) on train subset. 
```{r}
modFitgbm <- train(classe ~ ., method="gbm",data=subset_train,verbose=FALSE)
modFitgbm$finalModel
```

6. Show results
```{r}
predResultgbm <- predict(modFitgbm, subset_test)
confusionMatrix(table(subset_test$classe, predResultgbm))
```

Model accuracy is 
```{r}confusionMatrix(table(subset_test$classe, predResultgbm))[overall]$Accuracy```